\section{Future Development}
We acknowledge that there are features that may be nice to add to this tool in
the future. The authors of Instalog have considered the following features and
would like to add them, but currently believes they are infeasible due to time
constraints.
\begin{description}
\item[Embedded pevFind Engine] One of the authors has written another useful
tool called pevFind, which can enumerate useful information given the right
script. However, pevFind's architecture makes it infeasible to add at present.
\item[Self Defense] Kill known problematic processes at start in order to try to
run more reliably. Possibly work with Lawrence Abrams (also known as
``Grinler''), author of the RKill tool, to integrate its features.
\item[Rootkit Checking] Not possible currently as the authors have no kernel
development experience.
\item[Driver-based Deletions] Similar to the Avenger tool. Currently infeasible
as the authors have no kernel development experience.
\item[Batch Scripting] Ability to embed batch scripts into a script and have
them execute.
\item[Access Checking] Currently forums use the ``Junction'' tool for this,
which is not designed for this purpose, but ``works''.
\item[Differential Log Analysis] Allows a forum helper to analyze the
differences from one report to the next, instead of needing to analyze an entire
report at a time.
\item[Script File Restore] Would allow restoration of a backup from a script.
\item[WinRE Support] Would allow operation from the Windows Recovery
Environment.
\item[System Integrity Check] Checking of windows components against known
MD5s. Possibly in conjunction with rootkit checking feature.
\item[Reg Null Enumeration] Some kinds of malicious software will create
registry keys with NULL characters in their name to confuse C APIs. Some tools
enumerate this.
\end{description}
